---
title: Conditional Outputs with `blastula`
description: "A notebook that will send a customized emails using `blastula`"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor: visual
editor_options: 
  markdown:
    wrap: 72
    canonical: true
---

## Goals 

The goal of this document is to demonstrate different conditional
outcomes in rendering reports on Posit Connect:

-   Using `blastula` to create a different customized emails that will
    replace the default email sent by Posit Connect

-   Using `blastula` to suppress email output depending on condition

-   Using `knitr::knit_exit()` to stop executing code and exit the
    report rendering on condition

## Setup

### Load required packages for summarizing data and creating emails

```{r load-packages, message=FALSE, warning=FALSE}

library(tidyverse)
library(blastula)
library(glue)

```

## Generate Conditional Outcome

For demonstration, let's define 4 possible outcomes:

-   **case 1** - Send an email with a cat picture
-   **case 2** - Send an email with a dog picture
-   **case 3** - Do not send any emails but still render this notebook
    completely
-   **case 4** - Send a "something went wrong" email and prematurely end
    notebook rendering

```{r random-selection-of-case}

# randomly select a case 1, 2, 3, or 4
case <- sample(1:4, 1) 

```

For this rendering, it's going to be:

<h3>case `r case`!</h3>

## Render and attach the email (if appropriate)

With `blastula`, we first render the email, which is a separate Rmd
file. Because it's R Markdown, we can take advantage of markdown
formatting and mix prose with code.

```{r render-email}


if(case == 1){
  email <- blastula::render_connect_email("activity-04_case1_email.Rmd")
  subject <- glue("ðŸˆ Meow! Here's the cat of the day for {wday(today(), label = TRUE, abbr = FALSE)}")
}

if(case == 2){
  email <- blastula::render_connect_email("activity-04_case2_email.Rmd")
  subject <- glue("ðŸ• Woof! It's {wday(today(), label = TRUE, abbr = FALSE)}, here's a pup!")

}

if(case == 4){
  email <- blastula::render_connect_email("activity-04_case4_email.Rmd")
  subject <- glue("âš ï¸ There was an issue with the {wday(today(), label = TRUE, abbr = FALSE)} report")
}

if(case %in% c(1,2,4)){
  send_email <- TRUE
}else{send_email <- FALSE}

```

With the rendered email saved as the object `email`, we now attach it as
the output of this report using `blastula::attach_connect_email()`.

```{r attach-email}

# only do this if `send_email` is TRUE
if(send_email == TRUE){
  blastula::attach_connect_email(email,
                                 subject = subject)
# otherwise don't send any emails
}else{
  blastula::suppress_scheduled_email()
}

```

## Create the escape hatch

If it's case 4, we want to stop this report *right now*

```{r escape-hatch}

if(case == 4){knitr::knit_exit("ðŸ‘‹")}
```

Are you still there? Phew, that means it wasn't case 4 and we're still
going strong.

## Log what we did

We can have this report include the email output as part of its
rendering. This is great to be able to see what email was generated
during the rendering. Access historical renderings from the [Report
History option in
Connect](https://docs.posit.co/connect/user/report-history/).

Output the html of the rendered email by accessing the `html_html`
attribute of your rendered email object. Set the code chunk option
`results="asis"` to pass the html directly (R Markdown will not try to
process it)

Here's how it's done:

```{r add-logging, results='asis'}

# give a nice timestamp
if(case == 3){glue(
    "### Report run {blastula::add_readable_time()}

     ### This report was processed for **Case** {case}
     
     ### No email was sent.")}else
       {
glue("### Report run {blastula::add_readable_time()}

     ### This report was processed for **Case** {case}
     
     ### Here is a preview of the email rendered with this report: 
     
     ---")
     }

```

```{r add-email-render, results="asis" }

if(case %in% c(1,2)){
glue("**Subject: {subject}**")
email$html_html}

```
