---
title: "Session 1"
title-slide-attributes: 
  data-background-color: white
  data-background-image: _extensions/positconfslides/assets/backgrounds/content2-light.svg
  data-background-size: contain
format:
  positconfslides-revealjs: 
    chalkboard: true
    slide-number: c/t
    footer: "[http://pconf23-workflows.com/](http://pconf23-workflows.com/)"
    code-copy: true
    center-title-slide: false
    code-link: true
    code-overflow: wrap
    highlight-style: a11y
    width: "1600"
    height: "900"
    filters:
      - positconfslides
execute: 
  eval: true
  echo: true
  cache: true
---


## Workshop approach

We will use an [**end-to-end fictional project**]{.fuchsia} to demonstrate workflows and best practices using [**open source packages**]{.teal} and [**Posit professional tools**]{.teal}.

:::{.fragment}

<h2><br> Conventions </h2>

|                                               |                                                                            |
|-----------------------------------------------|----------------------------------------------------------------------------|
| 🧰                                            | Add this to your toolbox.                                                  |
| 📣                                            | I will stand on my soapbox and profess this until I am blue in the face.   |
| {{< bi sign-turn-right-fill color=orange >}}  | Detour warning. We could get really into this, but there's not time today. |
: {tbl-colwidths="[5,95]"}

:::





## About the Project

This workshop will use data from the City of Chicago Open Data Portal: <https://data.cityofchicago.org>.

::: smaller
1.  🍕 Food inspections: [https://data.cityofchicago.org/Health-Human-Services/Food-Inspections/4ijn-s7e5](https://data.cityofchicago.org/Health-Human-Services/Food-Inspections/4ijn-s7e5){.smaller}

```{r}
#| echo: false
head(pins::pin_read(pins::board_connect(), "katie.masiello/inspections_raw_test"),5) |> 
  dplyr::select(dba_name, license_number, facility_type, risk, address, results, violations) |> 
  dplyr::mutate(violations=stringr::str_trunc(violations,30))|> gt::gt()
```

2.  📒 Business licenses: [https://data.cityofchicago.org/Community-Economic-Development/Business-Licenses/r5kz-chrr](https://data.cityofchicago.org/Community-Economic-Development/Business-Licenses/r5kz-chrr){.smaller}

```{r}
#| echo: false
head(pins::pin_read(pins::board_connect(), "katie.masiello/bus_data_raw_test"),5) |>
  dplyr::select(license_number, doing_business_as_name, address, license_description, license_term_expiration_date) |> gt::gt()
```
:::

## Project Objective

-   Provide users with a self-service tool that predicts the likelihood of a facility failing their next inspection

<h2><br>Project Requirements</h2>

-   🤖 Automate the pipeline
-   ⚠️ Receive alerts if there are issues in the pipeline
-   🔄 Project is easy to maintain and iterate upon
-   ♻️ Work is reusable by other teams, even if they don't use R (Lookin' at you 🐍)

##  {.brackets-dark}

::: section-title-dark
Let's Get Started!
:::

## First Questions {.brackets-light}

::: incremental
-   What is this data?
-   Where should I put it?
:::




## Activity Highlights --- What is this Data?

:::{.callout-note icon=false}
## Activity
👉 Open the file `materials/activities/00_raw_data_exploration.qmd`
:::

Our first introduction to `pointblank`

::: columns
::: {.column width="15%"}
![](https://github.com/rstudio/pointblank/raw/main/man/figures/logo.svg){style="padding-right: 15px;"}
:::

::: {.column width="75%"}
<br>`pointblank` provides data quality assessment and metadata reporting for data frames and database tables. <https://github.com/rstudio/pointblank>
:::
:::


🧰 The `pointblank::scan_data()` function provides a HTML report of the input data to help you understand your data. 

## Sample data scan

```r
pointblank::scan_data(mtcars)
```
<br>


<iframe src="slide_resources/scan_mtcars.html" width=100% height=100%></iframe>
 
## Where should I put the data?

<br>

📣 Production data belongs in a database.

:::{.callout-tip icon=false}
## {{< bi sign-turn-right-fill color=orange >}} Best practices in working with databases
<https://solutions.posit.co/connections/db/> 

Specific topics:

- [Connect to a database](https://solutions.posit.co/connections/db/getting-started/connect-to-database/)
- [Query a database table](https://solutions.posit.co/connections/db/getting-started/database-queries/)
- [Securing credentials](https://solutions.posit.co/connections/db/best-practices/managing-credentials/)
- [Making scripts portable](https://solutions.posit.co/connections/db/best-practices/portable-code/)



:::
