---
title: "activity-02_publish_and_schedule_data_pull"
description: This document pulls our raw data and saves it to the database. Publish this to Connect and schedule it to run at an appropriate interval
toc: true
toc-depth: 4
editor: visual
editor_options: 
  markdown:
    wrap: 72
    canonical: true
  chunk_output_type: console
execute: 
  eval: false
---

## Activity Notes

This Notebook will simulate a real-life asset that sources raw data and
writes it to a database. We will publish this document to Connect and
schedule it to run at a regular interval, simulating a production
workflow. We'll add informational logs to our notebook that prove useful
when we review our published output and historical renderings on
Connect.

## Goals

The goals of this activity are to:

-   Gain experience using `DBI` commands interact with a database

-   Publish this notebook to Posit Connect and put it on a rendering
    schedule so database updates happen automatically

This will give you experience setting up a repeatable workflow for
populating production data sources.

✏️ There will be placeholders (`____`) in the code cells below that you
will need to fill in!

## Setup

### Load required packages for importing and writing the data

```{r}
#| label: Load packages

library(tidyverse)
library(janitor)
library(DBI)
library(RPostgres)
library(glue)

```

## Task 1 - Download raw data

🔄 Task

Just like the last activity, we will download the raw inspections data
from the Chicago Data Portal, but because this will become an automated
script, we will also include logging in our script that will provide
useful information when we review our published and scheduled output.

A useful informational log will be knowing how long the data took to
download. One means for doing this is to surround our download step with
time stamps, following this model:

``` r
start_time <- Sys.time()

#operation that we want to time

end_time <- Sys.time()
duration <- end_time - start_time
print(paste("Operation took", round(duration[[1]], 2),  units(duration)))
```

```{r}
#| label: Download raw inspections data from City of Chicago

## It's not necessary to run this chunk during the Activity time if you still have `inspections_raw` in memory from the last Activity. Save yourself some time 😉

# Start time stamp
start_time <- Sys.time()

inspections_raw <- 
  readr::read_csv(paste0(
    "https://data.cityofchicago.org/api/views/4ijn-s7e5/rows.csv?$$app_token=",
    Sys.getenv("SOCRATA_APP_TOKEN")
    ))

# End time stamp
end_time <- Sys.time()

print(paste("ℹ️ Info: Downloading Inspection data took", round(duration[[1]], 2),  units(duration)))

```

-   Use the `janitor` package to clean the column names so that they
    have no special characters or spaces, and they are the same case

-   Convert the data to uppercase for uniformity

Now we do a little cleaning of the data to make case consistent and
remove spaces and special characters from the column names.

It's a good idea to reformat the column names because databases are
sometimes unhappy with special characters.

```{r}
#| label: reformat column names

inspections_raw <- inspections_raw |> 
  #clean column names 
  janitor::____ |> 
  #convert data to all uppercase or lowercase for uniformity
  ____

head(inspections_raw)
```

## Task 2 - Write raw data to database

🔄 Task

-   Create a database connection using the database connection details
    defined for you by your IT Admin (Workshop Instructors)
-   Use `dbWriteTable` to write a subset of the `inspections_raw` data
    into the database (use a subset for sake of time) as
    `inspections_raw_<your_name>`

✅ Solution

The `RPostgres` package is DBI-compliant and is built specifically for
PosgreSQL databases. Performance may be faster (particularly for writing
data!) than using the generic `odbc` package.

```{r}
#| label: make database connection

# Using the `RPostgres` package  
con <- dbConnect(RPostgres::Postgres(), 
                 host = Sys.getenv("CONF23_DB_HOST"), 
                 port = "5432", 
                 dbname = "conf23_r", 
                 user = Sys.getenv("CONF23_DB_USER"), 
                 password = Sys.getenv("CONF23_DB_PASSWORD"))
con


```

For the sake of time, we don't want to write all \~260k rows of the
inspections data to the database, so first, we will take a small sample
and append our names to the data frame to provide unique names.

```{r}
#| label: make raw data subset

my_username <- ____

my_df_name <- paste0("inspections_raw_subset_", my_username)

# Take a subset of the data for the sake of reducing the time for processing. Reduce the row count under 5,000 to keep things snappy
inspections_raw_subset <- inspections_raw |> ____

inspections_raw_subset
```

Now we write the subset data to the database. We also use time stamps
for monitoring the write time, which will be useful information for our
scheduled reporting.

```{r}
#| label: database write


# write `inspections_raw_subset` with informational time stamps

# Insert your start time stamp
____

DBI::dbWriteTable(conn = ____, 
                  name = ____, 
                  value = ____,
                  overwrite = TRUE)


# Insert your end time stamp
____


print(paste("ℹ️ Info: Writing raw inspection data to database took", ____))
```

## Task 3 - Read from database

🔄 Task

Explore some typical database commands such as

-   `DBI::dbListTables()` to get a list of the tables in the database

-   `dplyr::tbl()` to interact with the database table you just wrote

-   Count the number of rows you wrote to the database in the table
    `inspections_raw_<your_name>`

✅ Solution

Use `DBI::dbListTables()` to get a list of the tables in the database,
and define database table with `dplyr::tbl()` to interact with
`inspections_raw_<your_name>`.

```{r}
#| label: list tables

DBI::dbListTables(____)

my_db_table <- dplyr::tbl(____, ____)
my_db_table
```

There are many ways to count database rows. However, if you just view
`my_db_table`, do you see this in the output:

```         
# Source:   table<inspections_raw_subset_katie.masiello> [?? x 17]
```

That's not very informative!

If we pull the database into local memory using `collect()`, we could
see (and count) all the rows, but we want to avoid this for performance
and resource sake. Try a `dplyr` aggregation command, like `tally`,
`count`, or `summarise` to give an answer without pulling the database
into memory.

```{r}
#| label: count table rows
# Be sure row_count is a numeric value, not a data frame (hint: use `dplyr::pull()` to extract a column as vector)
written_row_count <- my_db_table |> ____

#Close the database connection 
dbDisconnect(con)

```

## Task 4 - Add logging information to make our scheduled report informative

🔄 Task

Make your scheduled reports work for you! Include information that
you'll find useful to refer back to. Loggking can be as basic or
elaborate as you need.

Some suggestions include:

-   `blastula::add_readable_time()` 👈 we will come back to `blastula`
    in another activity, but this function is a nice one to add a
    friendly time stamp.

-   Summarize the main actions taken in the report, such as how many
    rows of data were downloaded, how many rows of data were written to
    the database

Hint: The `glue` package is helpful for combining text and variables. Or
just use `paste`.

✅ Solution

```{r}
#| label: logging

# Add some logging information here

```

🛑 Stop here for now. Put your "I'm good" sticky up to let us know
you're done with this Task and we will proceed to the publication step
momentarily.

## Task 5 - Publish and schedule on Posit Connect

🔄 Task

-   Use push-button publishing to publish this document, with source
    code, to Connect
-   Schedule it to run on a timescale that seems appropriate for the
    data update cycle

Notes: - Recall we specified our database connection details with
environment variables. Typically when you publish content to Connect,
you will also need to supply the environment variables using the "Vars"
pane in Connect. For this workshop, these environment variables have
already been stored on the Connect server for you so you won't need to
do this step for this activity.

✅ Solution

Let's do this task together as a group.

For reference, the Connect User Guide provides instructions for
publishing:
<https://docs.posit.co/connect/user/publishing/#publishing-general>
