---
title: "Shiny"
description: "Create a Shiny app that talks to the Food Inspections API"
toc: true
toc-depth: 4
editor_options: 
  markdown:
    canonical: true 
execute: 
  eval: false
editor: visual
---

## Setup

```{r}
#| label: load-packages

library(shiny)
library(tidyverse)
library(vetiver)
```

## Activity 1 - Review how to call Food Inspections API

In a moment, we will open up a Shiny application that will make interactive calls to the [Food Inspections API](https://connect.conf23workflows.training.posit.co/food-inspections-api/). Before we do that, let's review how to call the food inspections API by defining an endpoint, creating "new" data, and submitting a query to the API.

🔄 Tasks

-   Define the food inspections API endpoint.

-   Predict the likelihood that a food establishment will fail an inspections by querying the API (with mock data)

### Define API endpoint

Run the code below to define the endpoint of our API!

```{r}
# API URL
api_url <- "https://connect.conf23workflows.training.posit.co/food-inspections-api"

# Add predict
endpoint <- vetiver_endpoint(paste0(api_url, "/predict"))

endpoint
```

### Create mock data

Simulate a food establishment. Here is some information regarding the various features that will need to be added:

-   `facility_type`: string (example: "RESTAURANT")

-   `risk`: number (1 is high risk, 2 is medium risk, 3 is low risk)

-   `latitude`: number (example: 41.90)

-   `longitude`: number (example: -87.63)

-   `v_cumsum`: integer (cumulative number of violations - example: 30)

-   `v_cumsum_cs`: integer (cumulative number of critical and serious violations - example: 10)

-   `days_since_last_inspection`: number (example: 200)

```{r}
# Create mock data as a tibble
mock_establishment <- tibble(facility_type = "___",
                             risk = ___,
                             latitude = ___,
                             longitude = ___,
                             v_cumsum = ___,
                             v_cumsum_cs = ___,
                             days_since_last_inspection = ___)
```

Query the API using mock data:

```{r}
# Query the API and extract FAIL likliehood
predict(endpoint, mock_establishment)$.pred_FAIL
```
