---
title: "activity-03_data_validation.qmd"
description: "This notebook performs data validation using `pointblank`"
toc: true
toc-depth: 4
editor: visual
editor_options: 
  markdown:
    wrap: 72
    canonical: true
execute: 
  eval: false
---

## Goals

The goals of this activity are to:

-   Use the `pointblank` package to perform data validations

-   Define thresholds in data validation to be used for alerting

-   Create an informant data dictionary

✏️ There will be placeholders (`____`) in the code cells below that you
will need to fill in!

## Activity Notes

<!-- TODO: expand on this -->

This notebook is a simplified version of the deployed notebook used for
the project solution. This notebook is simplified to ????????
`01_data_clean_validate.rmd`

-   No business data
-   Just validation, no alerting yet
-   A few less validations
-   Skipped final mutations and cleanings that we do before training
    model
-   Omit summarization of data that is presented in an email
-   Omit write to database

Using filtered data that is cleaned up and filtered for only what is
relevant, meaning Facility Types of interest:

-   Restaurants

-   Grocery Stores

-   Bakeries

-   Coffee Shops

-   Wrigley Field Rooftop (because that sounds pretty fun)

We also will filter for only inspection entries where the results were:

-   Pass

-   Fail

-   Pass w/ conditions

## Setup

### Load required packages for data validation

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}

library(tidyverse)
library(pins)
library(pointblank)

```

### Read data from pin

✅ Solution

## [pointblank](https://github.com/rstudio/pointblank) is an R package that provides data quality assessment and metadata reporting for data frames and database tables.

![](https://github.com/rstudio/pointblank/raw/main/man/figures/logo.svg){fig-align="center"
width="212"}

The `pointblank::scan_data()` function provides a HTML report of the
input data to help you understand your data. It contains 6 sections:

-   

```{r read-data-from-pin}

board <- board_connect()

inspections_filtered <- board |> 
  pin_read("katie.masiello/inspections_filtered")
```

## Task 1 -

🔄 Task

Use the `pointblank` package to

-   create an agent
-   define validations
-   interrogate the table

✅ Solution

Activity 1 - First Validation of Data for Size, Shape, and Format of
Data

🔄 Task

Validate that:

-   the data has the expected columns
-   the column formats are as expected
-   there are at least 100k rows of data (an arbitrary value, but if
    it's less than this, we're missing a lot of data)

✅ Solution

This is our first validation. It just confirms the overall size and
format of the data frame. If the validation fails here, a warning is
thrown and the validation will not proceed.

```{r data-fram-size-and-format-validation}


# Define a column schema so we can check inputted data is as expected
# troubleshooting tip, if the validation fails, look at the x_list$col_types and $col_names to see the discrepancy
schema_inspections <- col_schema(inspection_id = "numeric",
                               dba_name = "character",
                               aka_name = "character",
                               license_number = "numeric",
                               facility_type = "character",
                               risk = "character",
                               address = "character",
                               city = "character",
                               state = "character",
                               zip = "numeric",
                               inspection_date = "Date",
                               inspection_type = "character",
                               results = "character",
                               violations = "character",
                               latitude = "numeric",
                               longitude = "numeric",
                               location = "character"
                               )


#### VALIDATION 1: Data set integrity validations. All of these trigger a warning under fail conditions.
agent_df_integrity <- 
  create_agent(inspections_filtered, 
               label = "Inital validation of the inspections data set to confirm overall schema and size. If there are issues with this validation, further processing stops and an alert is triggered.") |> 
  # verify column schema 
  col_schema_match(schema_inspections, 
                   label = "Is the column schema as expected?", 
                   actions = action_levels(warn_at = 1)) |> 
  #Check that expected columns exist. 
  # To do this, we modify the input df and call it a "precondition." The modification is made using a unique pointblank function `tt_tbl_colnames` that turns the column names of our input df into a data frame perfectly formatted to pointblank's preferences. That table is then compared to the set of schema_inspection names. See other cool transforms named `tt_*`
  col_vals_in_set(columns = value, 
                  set = names(schema_inspections), 
                  preconditions = ~. %>% tt_tbl_colnames, 
                  label = "Are the expected columns in the data set?", 
                  actions = action_levels(warn_at = 0.01) ) |> 
  # verify there are A LOT of rows of data to be sure import didn't mess up. 
  # If we didn't get a full set of data, we don't want to proceed. 
  col_vals_gte(columns = n, 
               value = 25000L, # an arbitrary high-ish number
               preconditions = ~. %>% tally,
               label = "Are there more than 25k rows in the data?", 
               actions = action_levels(warn_at = 1)) |>
  interrogate()
agent_df_integrity

# We'll make a variable `alert_me` which will be a logical TRUE/FALSE. We'll talk about *how* to alert in the next activity, so just hold this in the back of your mind for now.
alert_me <- !(all_passed(agent_df_integrity))
```

# Activity 2 - Second validation for data integrity

🔄 Task

Validate that the data itself

-   does not have null/NA values
-   has valid entries for Inspection Date,
-   has no repeated Inspection IDs
-   License Number is within a valid range
-   Latitude and Longitude are within a bounding box surrounding Chicago
-   Zip Code follows the expected standard

Bonus: Only run this validation if there was not an alert in the
previous validation

✅ Solution

Here's a logical statement we can use to stop processing this file if
the previous validation failed.

```{r optional-knit_exit-if-alert-is-true}
if (alert_me == TRUE){
  glue("Validation of inspection data set integrity failed. No additional validations were run, and processed data was NOT written to the database")}

# 🧰 stop processing this file
if (alert_me == TRUE){knitr::knit_exit()} 
# if alert_me is FALSE, all is good and let's go!
```

As long as the the validation on Data Frame integrity passes, the
following additional validations will run.

```{r data-integrity-validation}


#### VALIDATION 2: Data integrity validations. Failing rows from this validation will be stripped out. 
agent_main <-  
  create_agent(inspections_filtered, label = "Validation of inspections data") |> 
  # No null values
  col_vals_not_null(columns = inspection_id) |> 
  col_vals_not_null(columns = dba_name) |> 
  col_vals_not_null(columns = address) |> 
  col_vals_not_null(columns = latitude) |>
  col_vals_not_null(columns = longitude) |>
  col_vals_not_null(columns = inspection_date) |> 
  col_vals_not_null(columns = inspection_type) |> 
  col_vals_not_null(columns = results) |> 
  # verify inspection date is valid
  col_vals_lte(columns = inspection_date, today(),
               label = "Is inspection date in the past?") |> 
  # verify Inspection ID is unique
  rows_distinct(columns = vars(inspection_id), 
                label = "Is the Inspection ID unique?", 
                actions = action_levels(warn_at = 0.001)) |> 
  # verify license number is valid
  col_vals_between(columns = license_number, 1, 99999999, 
                   label = "Is the License Number a valid entry?") |>
  # verify lat and long bounds (set na_pass = TRUE to ignore NAs)
  col_vals_between(columns = latitude, left = 41.5001, right = 42.3772, na_pass = TRUE) |> 
  col_vals_between(columns = longitude, left = -88.2959, right = -87.316, na_pass = TRUE) |> 
  # verify w col_vals_within_spec for postal_code / aka zip
  col_vals_within_spec(columns = zip, spec = "postal[USA]", na_pass = TRUE) |> 
  interrogate() 

agent_main

```

# Activity 3 - Remove the failed data from the data set

🔄 Task

Pointblank has identified all of the rows that passed and failed
validation. Now remove those that failed so the data that is passed
downstream is squeaky clean.

✅ Solution We will "sunder" (sunder \| ˈsəndər \|verb \| split apart)
the data using `pointblank::get_sundered_data()`.

```{r sunder-data}

# Passed data
inspections_validated <- get_sundered_data(agent_main, type = "pass")

# Failed data
inspections_fail_validation <- get_sundered_data(agent_main, type = "fail")

```

# Activity 4 - Provide a data dictionary (aka Informant) to explain the data

🔄 Task

Data should never be used blindly. Proper documentation is essential.
Pointblank can generate an `informant` table that can provide both
static and dynamically-rendered details about your data.

-   Explore the table below.

✅ Solution

Table information can be synthesized in an information management
workflow, giving us a snapshot of a data table we care to collect
information on. The pointblank informant is fed a series of `info_*()`
functions to define bits of information about a table. This info text
can pertain to individual columns, the table as a whole, and whatever
additional information makes sense for your organization. We can even
glean little snippets of information (like column stats or sample
values) from the target table with info_snippet() and the `snip_*()`
functions and mix them into the data dictionary wherever they're needed.

Informant table / Data Dictionary

```{r data-dictionary}

create_informant(
  tbl = inspections_validated,
  label = "Data Dictionary - Validated Data from Chicago Food Inspection Dataset") |> 
  info_tabular(
    description = "This table defines the columns in the validated dataset from City of Chicago Food Inspection dataset"
  ) |> 
  info_section(
    section_name = "further information",
    `source` = "Data from City of Chicago, <https://www.cityofchicago.org>",
    `🔄 updates` = "Original data source updated every Friday. This table last updated {Sys.Date()}.",
    `ℹ️ note` = "This example material using data that has been modified for use from its original source, <https://www.cityofchicago.org>, the official website of the City of Chicago.  The City of Chicago makes no claims as to the content, accuracy, timeliness, or completeness of any of the data provided at this site.  The data provided at this site is subject to change at any time.  It is understood that the data provided at this site is being used at one's own risk."
  ) |> 
  info_columns(
    columns = inspection_id,
    `ℹ️` = "A unique record number."
  ) |> 
  info_columns(
    columns = dba_name,
    `ℹ️` = "'Doing Business As.' This is legal name of the establishment."
  ) |> 
  info_columns(
    columns = aka_name,
      `ℹ️` = "'Also Known As.' This is the name the public would know the establishment as."
  ) |> 
  info_columns(
    columns = license_number,
      `ℹ️` = "This is a unique number assigned to the establishment for the 
purposes of licensing by the Department of Business Affairs and Consumer Protection."
  ) |>  
  info_columns(
    columns = facility_type,
      `ℹ️` =
      "Cleaned and validated data includes only establishments described by one of the following: \n - **Bakery** \n - **Coffee Shop** \n - **Grocery Store** \n - **Restaurant** \n - **Wrigley Field Rooftop**"
  )  |>
  info_columns(
    columns = risk,
      `ℹ️` = "Risk category of facility. Each establishment is categorized as to its risk of adversely affecting the public's health, with 1 being the highest and 3 the lowest. The frequency of inspection is tied to this risk, with risk 1 establishments inspected most frequently and risk 3 least frequently."
  ) |>  
  info_columns(
    columns = address,
      `ℹ️` = "Street address, city, state and zip code of facility: This is the complete address where the facility is located."
  ) |>  
  info_columns(
    columns = city,
      `ℹ️` = "City of the facility."
  ) |>  
  info_columns(
    columns = state,
      `ℹ️` = "State of the facility."
  ) |>  
  info_columns(
    columns = zip,
      `ℹ️` = "Zip Code of the facility."
  ) |>  
  info_columns(
    columns = inspection_date,
      `ℹ️` = "This is the date the inspection occurred. A particular establishment is likely to have multiple inspections which are denoted by different inspection dates."
  ) |>  
  info_columns(
    columns = inspection_type,
      `ℹ️` = "An inspection can be one of the following types: \n - **canvass**: the most common type of inspection performed at a frequency relative to the risk of the establishment \n - **consultation**: when the inspection is done at the request of the owner prior to the opening of the establishment \n - **complaint**: when  the inspection is done in response to a complaint against the establishment \n - **license**: when the inspection is done as a requirement for the establishment to receive its license to operate \n - **suspect food poisoning**: 🤢when the inspection is done in response to one or more persons claiming to have gotten ill as a result of eating at the establishment (a specific type of complaint-based inspection) \n - **task-force inspection**: when an inspection of a bar or tavern is done. \n Re-inspections can occur for most types of these inspections and are indicated as such."
  ) |>  
  info_columns(
    columns = results,
      `ℹ️` = "An inspection can pass, pass with conditions or fail. Establishments receiving a 'pass' were found to have no critical or serious violations (violation number 1-14 and 15-29, respectively). Establishments receiving a 'pass with conditions' were found to have critical or serious violations, but these were corrected during the inspection. Establishments receiving a 'fail' were found to have critical or serious violations that were not correctable during the inspection. An establishment receiving a 'fail' does not necessarily mean the establishment's licensed is suspended."
  ) |>  
  info_columns(
    columns = violations,
      `ℹ️` = "An establishment can receive one or more of 45 distinct violations (violation numbers 1-44 and 70). For each violation number listed for a given establishment, the requirement the establishment must meet in order for it to NOT receive a violation is noted, followed by a specific description of the findings that caused the violation to be issued."
  ) |>  
  info_columns(
    columns = latitude,
      `ℹ️` = "The Latitude of the facility."
  ) |>
  info_columns(
    columns = longitude,
      `ℹ️` = "The Longitude of the facility."
  ) |>
  info_columns(
    columns = location,
      `ℹ️` = "The Latitude and Longitude of the facility."
  ) |>
  incorporate()


```
