---
title: "activity-01_publish_and_schedule_data_pull"
description: This document pulls our raw data and saves it to the database. Publish this to Connect and schedule it to run at an appropriate interval
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor: visual
editor_options: 
  markdown:
    wrap: 72
    canonical: true
---

## Goals

The goals of this activity are to:

-   Use a database connection to write our production data to our
    database

-   Gain experience using `DBI` commands to list database table
    information

-   Publish this notebook to Posit Connect and put it on a rendering
    schedule so database updates happen automatically

This will give you experience setting up a repeatable workflow for
populating production data sources.

✏️ There will be placeholders (`____`) in the code cells below that you
will need to fill in!

## Activity Notes

This notebook is a simplified version of the deployed notebook used for
the project solution. This notebook is stripped down to only pull the
inspection raw data and write a small subset to the database (business
data omitted just to keep things speedier). You may find it interesting
to browse the project notebook
(`materials/project/00_raw_data_scan_db_write.rmd`) to see additional
elements that make for a more robust component in the project pipeline.
The project solution has these additional features:

-   A check to ensure that at least 100k rows were written to the
    database. Failure of this would indicate an issue with the input
    data source.
-   An email alert if the number of rows written is not more than 100k
-   A pointblank data scan for reference and exploration
-   Ability to set the document to "TEST MODE" which will speed up
    processing by pulling a smaller test mode raw data set, avoids
    writes to the database, and generates the alert email for inspection

## Setup

### Load required packages for importing and exploring the data

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}

library(tidyverse)
library(janitor)
library(DBI)
library(RPostgres)
library(glue)

# Set eval=FALSE so incompleted activity will still render
# knitr::opts_chunk$set(eval = FALSE)
```

## Task 1 - Download raw data and add informational logs

🔄 Task

-   Just like the last activity, download the raw inspections data from
    the Chicago Data Portal

-   Use the `janitor` package to clean the column names so that they
    have no special characters or spaces, and they are the same case

-   Convert the data to uppercase for uniformity

-   Include informational logs in our script that will provide useful
    information when we review our published and scheduled output

✅ Solution

A useful informational log will be knowing how long the data took to
download. One means for doing this is to surround our download step with
time stamps, following this model:

``` r
start_time <- Sys.time()

#operation that we want to time

end_time <- Sys.time()
duration <- end_time - start_time
print(paste("Operation took", round(duration[[1]], 2),  units(duration)))
```

The `tictoc` package is another way to time events!

```{r Download-data-from-Chicago-Data-Portal}

# Insert your start time stamp
start_time <- Sys.time()

inspections_raw <- 
  readr::read_csv(paste0("https://data.cityofchicago.org/api/views/4ijn-s7e5/rows.csv?$$app_token=",
                         Sys.getenv("SOCRATA_APP_TOKEN")))

# Insert your end time stamp
end_time <- Sys.time()

duration <- end_time - start_time
print(paste("ℹ️ Info: Downloading Inspection data took", round(duration[[1]], 2),  units(duration)))

```

Now we do a little cleaning of the data to make case consistent and
remove spaces and special characters from the column names. It's a good
idea to reformat the column names because databases are sometimes
unhappy with special characters.

```{r}
inspections_raw <- inspections_raw |> 
  #clean column names 
  janitor::clean_names() |> 
  #convert data to uppercase for uniformity
  mutate(across(where(is.character), toupper))

head(inspections_raw)

```

## Task 2 - Write raw data to database

🔄 Task

-   Create a database connection using the database connection details
    defined for you by your IT Admin (Workshop Instructors)
-   Use `dbWriteTable` to write a subset of the `inspections_raw` data
    into the database (use a subset for sake of time) as
    `inspections_raw_<your_name>`

✅ Solution

The `RPostgres` package is DBI-compliant and is built specifically for
PosgreSQL databases. Performance may be faster (particularly for writing
data!) than using the generic `odbc` package.

```{r make-database-connection}

# Using the `RPostgres` package  
con <- dbConnect(RPostgres::Postgres(), 
                 host = Sys.getenv("CONF23_DB_HOST"), 
                 port = "5432", 
                 dbname = "conf23_r", 
                 user = Sys.getenv("CONF23_DB_USER"), 
                 password = Sys.getenv("CONF23_DB_PASSWORD"))
con

```

For the sake of time, we don't want to write all \~260k rows of the
inspections data to the database, so first, we will take a small sample
and append our names to the data frame to provide unique names.

```{r make-raw-data-subset}

my_username <- "katie.masiello"

my_df_name <- paste0("inspections_raw_subset_", my_username)

# Take a sample of only 1% of the data for the sake of reducing the time for processing (est. ~2.6k rows)
inspections_raw_subset <- inspections_raw |> sample_frac(0.01)

inspections_raw_subset
```

Now we write the subset data to the database. We also use time stamps
for monitoring the write time, which will be useful information for our
scheduled reporting.

```{r database-write}
# write `inspections_raw_subset` with informational time stamps

# Insert your start time stamp
start_time <- Sys.time()

DBI::dbWriteTable(conn = con, 
                  name = my_df_name, 
                  value = inspections_raw_subset,
                  overwrite = TRUE)


# Insert your end time stamp
end_time <- Sys.time()

duration <- end_time - start_time


print(paste("ℹ️ Info: Writing raw inspection data to database took", round(duration[[1]], 2),  units(duration)))

```

## Task 3 - Read from database

🔄 Task

Explore some typical database commands such as

-   `DBI::dbListTables()` to get a list of the tables in the database

-   `dplyr::tbl()` to interact with the database table you just wrote

-   Count the number of rows you wrote to the database in the table
    `inspections_raw_<your_name>`

✅ Solution

Use `DBI::dbListTables()` to get a list of the tables in the database,
and define database table with `dplyr::tbl()` to interact with
`inspections_raw_<your_name>`.

```{r list-tables}

DBI::dbListTables(con)

my_db_table <- dplyr::tbl(con, my_df_name)
my_db_table
```

There are many ways to count database rows. However, if you just view
`my_db_table`, do you see this in the output:

```         
# Source:   table<inspections_raw_subset_katie.masiello> [?? x 17]
```

That's not very informative!

If we pull the database into local memory using `collect()`, we could
see (and count) all the rows, but we want to avoid this for performance
and resource sake. Try a `dplyr` aggregation command, like `tally`,
`count`, or `summarise` to give an answer without pulling the database
into memory.

```{r count-table-rows}

# Be sure row_count is a numeric value, not a data frame (hint: use `dplyr::pull()` to extract a column as vector)
written_row_count <- my_db_table |> tally() |> pull()

#Close the database connection 
dbDisconnect(con)

```

## Task 4 - Add logging information to make our scheduled report informative

🔄 Task

Make your scheduled reports work for you! Include information that
you'll find useful to refer back to. Loggking can be as basic or
elaborate as you need.

Some suggestions include:

-   `blastula::add_readable_time()` 👈 we will come back to `blastula`
    in another activity, but this function is a nice one to add a
    friendly time stamp.

-   Summarize the main actions taken in the report, such as how many
    rows of data were downloaded, how many rows of data were written to
    the database

Hint: The `glue` package is helpful for combining text and variables. Or
just use `paste`.

✅ Solution

```{r logging}

# Add some logging information here

glue("Report run {blastula::add_readable_time()}")

glue("{nrow(inspections_raw)} rows of inspection data downloaded. 
     {written_row_count} rows of inspection data written to database.")


```

🛑 Stop here for now. Put your "I'm good" sticky up to let us know
you're done with this Task and we will proceed to the database section
momentarily.

## Task 5 - Publish and schedule on Posit Connect

🔄 Task

-   Use push-button publishing to publish this document, with source
    code, to Connect
-   Schedule it to run on a timescale that seems appropriate for the
    data update cycle

Notes: - Recall we specified our database connection details with
environment variables. Typically when you publish content to Connect,
you will also need to supply the environment variables using the "Vars"
pane in Connect. For this workshop, these environment variables have
already been stored on the Connect server for you so you won't need to
do this step for this activity.

✅ Solution

Let's do this task together as a group.

For reference, the Connect User Guide provides instructions for
publishing:
<https://docs.posit.co/connect/user/publishing/#publishing-general>
