---
title: "Raw Data Exploration"
description: "Conduct ad hoc exploration of the data, first exposure to `pointblank`"
toc: true
toc-depth: 4
editor: visual
editor_options: 
  markdown:
    wrap: 72
    canonical: true
  chunk_output_type: console
execute: 
  eval: false
---

------------------------------------------------------------------------

## General Workshop Activity Notes

Welcome to the first Activity! A few points relevant to this, and all
other Activities in the workshop:

1.  The Activities are either R Markdown or Quarto documents, presented
    by default in the Visual Editor for readability
2.  The intention is to run through the document from top to bottom,
    running each code chunk interactively rather than rendering/knitting
    the document
3.  Run the chunk by either clicking the green Play Button in the
    chunk's top right corner, or use the cmd+enter keyboard shortcut to
    run the current line or selected text
4.  We'll work through these activities mostly as a group, code-along
    style because our focus in this workshop is less about code
    development and more about around processes, tools, and learning
    concepts. We can best teach this if we do it together!
5.  **Where's my output?** Are you a `Chunk Output Inline` kind of
    person? Sorry. For this activity, it's set by default to go the
    console because the HTML renderings from Pointblank get very
    squished otherwise. It might be like this for other activities too,
    so don't get surprised 😉

------------------------------------------------------------------------

## Goals

The goals of this activity are to:

-   get familiar with the Food Inspections data set, which is used
    throughout this Workshop
-   gain experience using `pointblank::scan_data` to summarize data

This will give you the first clues for what data cleaning and
validations will be relevant for this project.

✏️ There will be placeholders (`____`) in the code cells below that you
will need to fill in!

## Setup

This workshop will use data from the City of Chicago Open Data Portal:
<https://data.cityofchicago.org>.

For this activity, we will explore the Food Inspections data set:

<https://data.cityofchicago.org/Health-Human-Services/Food-Inspections/4ijn-s7e5>

### Load required packages for importing and exploring the data

```{r}
#| label: Load packages

library(tidyverse)
library(pointblank)
library(ggforce) 
library(tidyverse)

```

### Download raw data

The City of Chicago data portal provides access to their data sets via
the Socrata Open Data (SODA) API.

We will query the SODA API directly and download the full data set as a
`.csv` using `readr::read_csv()`.

Note we are using an environment variable called **SOCRATA_APP_TOKEN**.
This has been preconfigured for you for the purposes of this workshop.

```{r}
#| label: Download raw inspections data from City of Chicago

# run this code as-is
inspections_raw <- 
  readr::read_csv(paste0(
    "https://data.cityofchicago.org/api/views/4ijn-s7e5/rows.csv?$$app_token=",
    Sys.getenv("SOCRATA_APP_TOKEN")
    ))

inspections_raw
```

Maybe you're already familiar with `str`, `summary` for preliminary data
exploration. Do you know `skimr::skim()`?

```{r}
str(inspections_raw)
```

```{r}
summary(inspections_raw)
```

```{r}
skimr::skim(inspections_raw)
```

## Task 1 - Scan the data with \`pointblank::scan_data()\`

🔄 Task

Use the `pointblank` package to gather basic information about the Food
Inspections data so we can

-   Understand what the data is
-   Get a sense for how much cleaning is required
-   Plan our approach for data validation

The `pointblank::scan_data()` function provides a HTML report of the
input data to help you understand your data. It contains 6 sections:

-   **Overview (O)**: Table dimensions, duplicate row counts, column
    types, and reproducibility information
-   **Variables (V)**: A summary for each table variable and further
    statistics and summaries depending on the variable type
-   **Interactions (I)**: A matrix plot that shows interactions between
    variables
-   **Correlations (C)**: A set of correlation matrix plots for
    numerical variables
-   **Missing Values (M)**: A summary figure that shows the degree of
    missingness across variables
-   **Sample (S)**: A table that provides the head and tail rows of the
    data set

The scan can take a little while to run on a large data set, but you can
also omit sections that are not needed.

First run this example on a small data frame, such as `penguins` from
the `palmerpenguins` package 🐧:

```{r}
#| label: Data scan on penguins

pointblank::scan_data(palmerpenguins::penguins) 

```

⚠️ The "Correlations" and "Interactions" sections take the longest to
run, especially on a larger data set like our Food Inspection data.

Run the `data_scan` on `inpsections_raw`, but omit "Correlations (C)"
and "Interactions (I)" because they are time consuming and not relevant
for this data.

```{r}
#| label: Data scan on inspections_raw

scan <- pointblank::scan_data(inspections_raw, sections = "OVMS")
scan

```

## Task 2 - Explore the data scan

🔄 Task

Let's explore the data scan 🔍

1.  What is the most common value for `DBA NAME` and `AKA Name` (hint:
    Toggle details)
2.  What `License #` has the most inspections?
3.  Explore the `Facility Type`. Does the number of distinct values
    surprise you? (the description of the data on
    <https://data.cityofchicago.org> says there are 23 categories!)
4.  What can we the Missing Values diagram tell us about potential data
    quality issues?

How will this inform our data cleaning and validation?
