---
title: "activity-03_data_validation.qmd"
description: "This notebook performs data validation using `pointblank`"
toc: true
toc-depth: 4
editor: visual
editor_options: 
  markdown:
    wrap: 72
    canonical: true
execute: 
  eval: true
---

## Goals

The goals of this activity are to:

-   Use the `pointblank` package to perform data validations

-   Define thresholds in data validation to be used for alerting

-   Create an informant data dictionary

✏️ There will be placeholders (`____`) in the code cells below that you
will need to fill in!

## Setup

### Load required packages for data validation

```{r}
#| label: Load packages

library(tidyverse)
library(pointblank)

```

## Task 1 - Data validation on a simple table

🔄 Task

Use the `pointblank` package to validate a simple table, provided by
`pointblank::small_table`

-   create an agent
-   define validations
-   interrogate the table

✅ Solution

Let's inspect `small_table`. We see a few different types of columns.
Rows 9 and 10 are exact duplicates. The `c` column contains two `NA`
values. This will serve us well for some validation exercises.

```{r}
#| label: inspect small_table

small_table

```

First we create an **agent.**

```{r}
#| label: create agent for small_table

agent <- pointblank::create_agent(small_table)
agent
```

On it's own, the agent is not very informative. It's waiting for
validations to be defined and an interrogation action to be performed.

Now we define our **data validation functions**. A few have been started
for you as examples. Create your own and explore! Refer to the package
documentation for the validation function reference:
<https://rstudio.github.io/pointblank/reference/index.html#validation-expectation-and-test-functions>

```{r}
#| label: define validations for small_table

agent <- agent |> 
  # are column `a` values between 1 and 5?
  col_vals_between(a, 1, 5) |> 
  # are column `f` values in the set of c("high", "med", "low")
  col_vals_in_set(f, c("high", "med", "low")) |> 
  # Now add your own!
  # are there any NA or NULL values
  col_vals_not_null(vars(a, b, c, d, e, f))


agent

```

If we look at the output of `agent`, it shows our validation plan, but
the action is yet to come when we **interrogate**.

```{r}
#| label: interrogate the small_table agent

agent |> interrogate()

```

Explore the validation report. Can you:

1.  Identify what fail percentage each validation had?
2.  Identify how many rows failed each validation?
3.  View the CSV extracts of failed data?

## Task 2 - Add Action Levels

🔄 Task

Iterate on your agent created for `small_table` and add action levels.
Action levels behave like tags. You can decide what threshold you want
to put for `notify`, `warn`, and `stop`. At a minimum, the tag will
provide a visual indicator of threshold exceedance in the validation
table. You can also use these tags post-interrogation to take actions.

✅ Solution

The action levels can be set as an **integer**, representing the
threshold number of failing units , or a **fraction**, representing the
fraction of failing units.

Use
`actions = action_levels(warn_at = ____, stop_at = ____, notify_at = ____)`
to add action levels to one, some, or all of your validations and rerun
the interrogation to see the output. Some samples have been provided.

```{r}
#| label: validation of small_table with action levels

agent_with_actions <- create_agent(small_table) |> 
  # are column `a` values between 1 and 5?
  col_vals_between(a, 1, 5,
                   actions = action_levels(notify_at = 0.10)) |> 
  # are column `f` values in the set of c("high", "med", "low")
  col_vals_in_set(f, c("high", "med", "low"),
                  actions = action_levels(warn_at = 0.1, stop_at = 0.3)) |> 
  # Now add your own!
  # are there any NA or NULL values
  col_vals_not_null(vars(a, b, c, d, e, f),
                    actions = action_levels(warn_at = 0.2)) |> 
  
  interrogate()

agent_with_actions

```

## Task 3 - Remove failing data from the dataset

🔄 Task

The intel provided in `agent_with_actions` tells us all of the rows of
`small_table` that passed and failed validation. Now remove those that
failed so the data that is passed downstream is squeaky clean.

✅ Solution

`pointblank` provides a number of [post-interrogation
functions](https://rstudio.github.io/pointblank/reference/index.html#post-interrogation)
to work with intel gathered from the validation. For this task, we will
"sunder" the data using `pointblank::get_sundered_data()`.

> **💡 sunder** /sun·der / ˈsən-dər / *verb* \| to split apart

```{r}
#| label: sunder data

# Passed data
small_table_validated <- get_sundered_data(agent = agent_with_actions,
                                           type = "pass")

# Failed data
small_table_failed <- get_sundered_data(agent = agent_with_actions,
                                        type = "fail")

```

## Task 4 - Provide a data dictionary (aka Informant) to explain the data

🔄 Task

Data should never be used blindly. Proper documentation is essential.
Pointblank can generate an `informant` table (OR or data dictionary)
that can provide both static and dynamically-rendered details about your
data.

Create a data dictionary for `small_table_validated`.

✅ Solution

Table information can be synthesized in an information management
workflow, giving us a snapshot of a data table we care to collect
information on. The pointblank informant is fed a series of `info_*()`
functions to define bits of information about a table. This info text
can pertain to individual columns, the table as a whole, and whatever
additional information makes sense for your organization. We can even
glean little snippets of information (like column stats or sample
values) from the target table with `info_snippet()` and the `snip_*()`
functions and mix them into the data dictionary wherever they're needed.

First let's see the bare bones informant table that `pointblank` can
create.

```{r}
#| label: basic informant for small_table_validated

create_informant(
  tbl = small_table_validated
)

```

The bare table already gives a summary of the table size, shape, and
column schema. But Now let's add to this with some `info_*()` functions
to make it more useful. Reference the Informative Functions section of
the `pointblank` documentation:
<https://rstudio.github.io/pointblank/reference/index.html#information-functions>.
A few sections have been provided as examples. Notice that markdown
syntax is supported!

```{r}
#| label: create data dictionary 

create_informant(
  tbl = small_table_validated) |> 
  info_tabular(
    description = "Data Dictionary for validated `small_table` data"
    ) |> 
  info_section(
    section_name = "further information",
    `source` = "Data set provided as part of the `pointblank` package."
    ) |> 
  info_columns(
    columns = date_time,
    `Info:` = "The date and time of the **observation.**",
    # Let's add a calculated value. This `latest_date` value is defined in the snippet in the next section
    `Latest Date:` = "The latest date is {latest_date}."
    ) |> 
  info_snippet(
    snippet_name = "latest_date",
    # caution: pointblank hasn't learned the native pipe yet in internal functions Use %>%.
    fn = ~ . %>% dplyr::pull(date_time) %>% max(na.rm = TRUE)
    ) |>
  # Add your own information to make this table more meaningful.
  info_columns(
    columns = vars(a, c, d),
    `Info:` = "the number of times an attempt was made."
  ) |> 
  # Using `incorporate` will run any calculations defined within the table
  incorporate()


```
